---
title: "Get absolute quantification data"
author:
- name: Xiaotao Shen PhD (https://www.shenxt.info/)
- affiliation: School of Medicine, Stanford University
  name: Chuchu wang PhD
  url: http://shenxt.info/
date: "Created on 2021-02-09 and updated on `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: no
  pdf_document:
    toc: no
vignette: >
  %\VignetteIndexEntry{absolute_quantification}
  %\VignettePackage{metID}
  % \VignetteEngine{knitr::rmarkdown}
  % \usepackage[utf8]{inputenc}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
```

***

This step is used to get the absolute quantification data of all lipids in all samples. If you have not finished the previous step, please click here: [Get relative quantification data](). 

If you have finished this step: click here to next step: [Output results]().

# **Get absolute quantification**

We should have finished step 1 and step 2. 

Then we should read the relative quantification data of internal standards and lipids.

## **Positive mode**

```{r, eval=TRUE, warning=FALSE, message=FALSE, cache=TRUE}
library(lipidflow)
library(tidyverse)
library(openxlsx)
```

```{r, eval=TRUE, warning=FALSE, message=FALSE, cache=TRUE}
is_quantification_table =
  readxl::read_xlsx("example/POS/is_relative_quantification/is_quantification_table.xlsx")

lipid_quantification_table =
  readxl::read_xlsx("example/POS/lipid_relative_quantification/lipid_quantification_table.xlsx")
    
```

We also need to get the sample information.

```{r, eval=TRUE, warning=FALSE, message=TRUE, cache=TRUE}
pos_mzxml = list.files(
  path = "example/POS",
  pattern = "mzXML",
  all.files = TRUE,
  recursive = TRUE
)

sample_info_pos =
  pos_mzxml %>%
  stringr::str_split(pattern = "\\/") %>%
  do.call(rbind, .) %>%
  as.data.frame() %>% 
  dplyr::rename(group = V1, sample.name = V2) %>%
  dplyr::select(sample.name, group) %>% 
  dplyr::mutate(sample.name = stringr::str_replace(sample.name, "\\.mzXML", ""))
```

Then the `match_item` should be set as:

```{r, eval=TRUE, warning=FALSE, message=TRUE, cache=TRUE}
match_item_pos =
  list(
    "Cer" = "d18:1 (d7)-15:0 Cer",
    "ChE" = c("18:1(d7) Chol Ester", "Cholesterol (d7)"),
    "Chol" = "Cholesterol (d7)",
    "DG" = "15:0-18:1(d7) DAG",
    "LPC" = "18:1(d7) Lyso PC",
    "LPE" = "18:1(d7) Lyso PE",
    "MG" = "18:1 (d7) MG",
    "PA" = "15:0-18:1(d7) PA (Na Salt)",
    "PC" = "15:0-18:1(d7) PC",
    "PE" = "15:0-18:1(d7) PE",
    "PG" = "15:0-18:1(d7) PG (Na Salt)",
    "PI" = "15:0-18:1(d7) PI (NH4 Salt)",
    "PPE" = "C18(Plasm)-18:1(d9) PE",
    "PS" = "15:0-18:1(d7) PS (Na Salt)",
    "SM" = "d18:1-18:1(d9) SM",
    "TG" = "15:0-18:1(d7)-15:0 TAG"
  )
```

Then run `get_absolute_quantification()` function to get the absolute quantification data of positive mode.

```{r, eval=TRUE, warning=FALSE, message=TRUE, cache=TRUE}
absolute_data_pos = get_absolute_quantification(
  path = "example/POS/",
  is_quantification_table = is_quantification_table,
  lipid_quantification_table = lipid_quantification_table,
  sample_info = sample_info_pos,
  match_item = match_item_pos
)
```

Now all the absolute quantification results are outputted in `example/POS/absolute/quantification` folder.
    
    
## **Negative mode**

```{r, eval=TRUE, warning=FALSE, message=FALSE, cache=TRUE}
is_quantification_table =
  readxl::read_xlsx("example/NEG/is_relative_quantification/is_quantification_table.xlsx")

lipid_quantification_table =
  readxl::read_xlsx("example/NEG/lipid_relative_quantification/lipid_quantification_table.xlsx")
    
```

```{r, eval=TRUE, warning=FALSE, message=TRUE, cache=TRUE}
neg_mzxml = list.files(
  path = "example/NEG",
  pattern = "mzXML",
  all.files = TRUE,
  recursive = TRUE
)

sample_info_neg =
  neg_mzxml %>%
  stringr::str_split(pattern = "\\/") %>%
  do.call(rbind, .) %>%
  as.data.frame() %>% 
  dplyr::rename(group = V1, sample.name = V2) %>%
  dplyr::select(sample.name, group) %>% 
  dplyr::mutate(sample.name = stringr::str_replace(sample.name, "\\.mzXML", ""))
```

Then the `match_item` should be set as:

```{r, eval=TRUE, warning=FALSE, message=TRUE, cache=TRUE}
match_item_neg =
  list(
    "Cer" = "d18:1 (d7)-15:0 Cer",
    "Chol" = "Cholesterol (d7)",
    "ChE" = c("18:1(d7) Chol Ester", "Cholesterol (d7)"),
    "LPC" = "18:1(d7) Lyso PC",
    "LPE" = "18:1(d7) Lyso PE",
    "PC" = "15:0-18:1(d7) PC",
    "PE" = "15:0-18:1(d7) PE",
    "PG" = "15:0-18:1(d7) PG (Na Salt)",
    "PI" = "15:0-18:1(d7) PI (NH4 Salt)",
    "PPE" = "C18(Plasm)-18:1(d9) PE",
    "PS" = "15:0-18:1(d7) PS (Na Salt)",
    "SM" = "d18:1-18:1(d9) SM"
  )
```

Then run `get_absolute_quantification()` function to get the absolute quantification data of negative mode.

```{r, eval=TRUE, warning=FALSE, message=TRUE, cache=TRUE}
absolute_data_neg = get_absolute_quantification(
  path = "example/NEG",
  is_quantification_table = is_quantification_table,
  lipid_quantification_table = lipid_quantification_table,
  sample_info = sample_info_neg,
  match_item = match_item_neg
)
```

Now all the absolute quantification results are outputted in `example/NEG/absolute/quantification` folder.
    
## **Combine positive and negative mode together**

Next, we need to combine positive and negative mode absolute quantification data together.

```{r, eval=TRUE, warning=FALSE, message=TRUE, cache=TRUE}
combine_pos_neg_quantification(
  path = "example/Result",
  express_data_abs_ug_ml_pos = absolute_data_pos$express_data_abs_ug_ml,
  express_data_abs_um_pos = absolute_data_pos$express_data_abs_um,
  variable_info_abs_pos = absolute_data_pos$variable_info_abs,
  express_data_abs_ug_ml_neg = absolute_data_neg$express_data_abs_ug_ml,
  express_data_abs_um_neg = absolute_data_neg$express_data_abs_um,
  variable_info_abs_neg = absolute_data_neg$variable_info_abs
)
```

Then all the results are outputted into the `example/Result` folder.


    
## **Next step**

Next we need to output some results (plots). Please click here: [Output results](https://jaspershen.github.io/lipidflow/articles/output_result.html)